New and simplified manual controls for projection and slice tours, with application to exploring classification boundaries in high dimensions

### 2022-09-22 paper/paper.Rmd
Iteration on the software section

The implementation of the manual tour with slice control needs a visualization of the current projection in terms of an axis display. This display should be able to track the mouse position and adjust the projection based on the user selection. The user can select one axis (corresponding to one variable) via clicking on it, and then adjust the position of that axis by dragging it during the click event. In practice the closest axis will be selected, and the dragging results in a tracking of the mouse position that will update the projection in small steps, rendering interpolation unnecessary.

The interaction in the axis display need to be mapped back onto the projection matrix, which in turn needs to be linked to both the orthonormalization procedure and to a second display that is showing the projected data.

@@

The most important inbuilt Mathematica function for our purpose is the locator pane. This creates a region on the screen where the position of the mouse is converted to input for the graphics functions, enabling the manual navigation of the tour.

\textcolor{red}{To Ursula:I removed the comment on the slider. There are no problems with color mapping. What do you mean by:" Also there seem to be a lot of loops needed?"? as far as I can tell the code does not use any loop. Is my explanantion of the locator pane enough? Please modify as you see fit and finish up this section.}

The examples presented in the penguins’ notebooks all use our primary new function, `SliceDynamic`. This function typically accepts grouped data in the form of a matrix where the second last column details the name of the group and the last column details the group index \textcolor{red}{(which can be one if there is only one group)}. After specifying the initial slice thickness and the slice range, the user is presented with an interactive display in which the control objects appear on the left and the slice visulisation appears on the right. The user can change the orientation of the slice via the locator pane, which changes the projection matrix;  a slider controls the slice thickness; and there is an input field, which changes the slice centre. The user can also change the appearance of the plot by zooming into the center or changing the point size with sliders provided. \textcolor{red}{It is worth noting that this zoom works best with data that has been centred and scaled.} The projected data can be displayed to contrast it with the slice by ticking a box. The explicit projection matrix can also be displayed via another checkbox. This is especially useful when we wish to import a projection that was identified as interesting in the manual exploration later, as shown in the applications below.

\textcolor{red}{Details about the implementation and usage instructions are given in the Appendix.}

@@

This short note has described new technology for manually interacting with tours. The manual control of projection coefficients is most useful for assessing variable importance to perceived structure, but can be generally used for steering a viewer through high-dimensional space. Changing the center of a slice manually enables exploring the space orthogonal to a projection, specifically in the direction of a single variable. This is a new tool, and as shown in the example, when used with the slice tour can very useful for understanding boundaries of classifiers. This is also likely a useful method for examining low-dimensional non-linear manifolds, and also functions of multiple parameters.

>>>>

The implementation of the manual tour as suggested here requires the visualization of the current projection in terms of an axis display (see Figure \ref{fig:manualsequence} for an example). This display should be able to track the mouse position and adjust the projection based on the user selection. The user can select one axis (corresponding to one variable) by clicking on it, and then adjust the position of that axis by dragging it during the click event. In practice the closest axis will be selected, and the dragging results in tracking of the mouse position which will update the projection in small steps, such that no further interpolation is required.

The interactions in the axis display need to be mapped back onto the current projection matrix, which will then be orthonormalized before feeding back into the axis display. A second, linked display shows the projected data in sync with the updates from the axis display.

@@

The most relevant inbuilt Mathematica function for our purpose is the locator pane. This creates a region on the screen where the position of the mouse is captured and then converted to input that updates the graphics functions, enabling the manual navigation of the tour.

The examples presented in the application section below (and attached in notebook format in the supplementary material) all use our primary new function, `SliceDynamic`. This function typically accepts grouped data in the form of a matrix where the second last column details the name of the group and the last column details the group index (which can be one if there is only one group). After specifying the initial slice thickness and the slice range, the user is presented with an interactive display in which the control objects appear on the left and the slice visulisation appears on the right. The user can change the orientation of the slice via the locator pane, which changes the projection matrix;  a slider controls the slice thickness; and there is an input field, which changes the slice centre. The user can also change the appearance of the plot by zooming into the center or changing the point size with sliders provided. It is worth noting that this zoom works best with data that has been centred and scaled. The projected data can be displayed to contrast it with the slice by ticking a box. The explicit projection matrix can also be displayed via another checkbox. This is especially useful when we wish to import a projection that was identified as interesting in the manual exploration later, as shown in the applications below.

Details about the implementation and usage instructions are given in the Appendix.

@@

This short note has described new technology for manually interacting with tours. The manual control of projection coefficients is most useful for assessing variable importance to perceived structure, but can be generally used for steering a viewer through high-dimensional space. Changing the center of a slice manually enables exploring the space orthogonal to a projection, specifically in the direction of a single variable. This is a new tool, and as shown in the example, when used with the slice tour can be very useful for understanding boundaries of classifiers. This is also likely a useful method for examining low-dimensional non-linear manifolds, and also functions of multiple parameters.

### 2022-09-22 paper/paper.tex
Iteration on the software section

The most important inbuilt Mathematica function for our purpose is the
locator pane. This creates a region on the screen where the position of
the mouse is converted to input for the graphics functions, enabling the
manual navigation of the tour.

\textcolor{red}{To Ursula:I removed the comment on the slider. There are no problems with color mapping. What do you mean by:" Also there seem to be a lot of loops needed?"? as far as I can tell the code does not use any loop. Is my explanantion of the locator pane enough? Please modify as you see fit and finish up this section.}

The examples presented in the penguins' notebooks all use our primary
new function, \texttt{SliceDynamic}. This function typically accepts
grouped data in the form of a matrix where the second last column
details the name of the group and the last column details the group
index \textcolor{red}{(which can be one if there is only one group)}.
After specifying the initial slice thickness and the slice range, the
user is presented with an interactive display in which the control
objects appear on the left and the slice visulisation appears on the
right. The user can change the orientation of the slice via the locator
pane, which changes the projection matrix; a slider controls the slice
thickness; and there is an input field, which changes the slice centre.
The user can also change the appearance of the plot by zooming into the
center or changing the point size with sliders provided.
\textcolor{red}{It is worth noting that this zoom works best with data that has been centred and scaled.}
The projected data can be displayed to contrast it with the slice by
ticking a box. The explicit projection matrix can also be displayed via
another checkbox. This is especially useful when we wish to import a
projection that was identified as interesting in the manual exploration
later, as shown in the applications below.

\textcolor{red}{Details about the implementation and usage instructions are given in the Appendix.}

>>>>

The most relevant inbuilt Mathematica function for our purpose is the
locator pane. This creates a region on the screen where the position of
the mouse is captured and then converted to input that updates the
graphics functions, enabling the manual navigation of the tour.

The examples presented in the application section below (and attached in
notebook format in the supplementary material) all use our primary new
function, \texttt{SliceDynamic}. This function typically accepts grouped
data in the form of a matrix where the second last column details the
name of the group and the last column details the group index (which can
be one if there is only one group). After specifying the initial slice
thickness and the slice range, the user is presented with an interactive
display in which the control objects appear on the left and the slice
visulisation appears on the right. The user can change the orientation
of the slice via the locator pane, which changes the projection matrix;
a slider controls the slice thickness; and there is an input field,
which changes the slice centre. The user can also change the appearance
of the plot by zooming into the center or changing the point size with
sliders provided. It is worth noting that this zoom works best with data
that has been centred and scaled. The projected data can be displayed to
contrast it with the slice by ticking a box. The explicit projection
matrix can also be displayed via another checkbox. This is especially
useful when we wish to import a projection that was identified as
interesting in the manual exploration later, as shown in the
applications below.

Details about the implementation and usage instructions are given in the
Appendix.
